---
title: "Figures_peers"
author: "Camila Scaff"
date: "5/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(scales)
library(gridExtra)
library(ggpubr)
library(reshape2)
library(dplyr)

#read the right tables
read.csv("results_Rmd/demuth_speakers.txt", sep = "") -> sum_demuth_tot #contains unclassified speech and investigator
read.delim("results_Rmd/french_speakers.txt", sep = "") -> sum_french_tot #contains unclassified speech and investigator
read.delim("results_Rmd/french_register.txt", sep = "") -> tot_sum_french_register # contains unclassified speech but only agreement between coders
read.delim("results_Rmd/demuth_register.txt", sep = "") -> tot_sum_demuth_register# contains unclassified speech but only agreement between coders
table(tot_sum_demuth_register$keychild,tot_sum_demuth_register$speaker) #check
table(tot_sum_demuth_register$keychild,tot_sum_demuth_register$input) #check

#create the lines (this should be done via a loop sorry!)
french_fix <- data.frame(
  keychild=c("Anais", "Theotime",  "Theotime",  "Theotime", "Theotime"),
  input=c("OCDS", "OCDS","ADS", "CDS","OCDS"),
  speaker=c("adult","adult","adult", "adult", "other child"),
  nb_utt=c(0,0,0,0,0))

#added to the main data.frame
tot_sum_french_register <- rbind(tot_sum_french_register, french_fix)

tot_sum_french_register$nb_utt <- as.numeric(tot_sum_french_register$nb_utt)

```

```{r functions, echo=F}
#defining multiplot to create easy panels
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  if (numPlots==1) {
    print(plots[[1]])
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=14, face="bold")
  )

data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}

summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}
```


Snap Shot of the different speakers in the corpora
```{r hundredpercent, echo=F}
##determine 100%
total_demuth<-sum(sum_demuth_tot$nb_utt) #69573
total_french<-sum(sum_french_tot$nb_utt) #37843

## determine 100% of speakers REMOVING keychild production
tot_speaker_demuth<-sum(sum_demuth_tot$nb_utt[sum_demuth_tot$speaker== "Target_Child"]) #28853
tot_speaker_demuth*100/total_demuth -> d_perc_TC #41.5%

tot_speaker_french<-sum(sum_french_tot$nb_utt[sum_french_tot$speaker== "Target_Child"]) #16338
tot_speaker_french*100/total_french -> fr_perc_TC #42.6%

inv_speaker_demuth<-sum(sum_demuth_tot$nb_utt[sum_demuth_tot$speaker== "investigator"]) #5486
inv_speaker_demuth*100/total_demuth -> d_perc_INV #7.9%

inv_speaker_french<-sum(sum_french_tot$nb_utt[sum_french_tot$speaker== "investigator"]) #2333
inv_speaker_french*100/total_french -> fr_perc_INV #6.1%

a_speaker_demuth<-sum(sum_demuth_tot$nb_utt[sum_demuth_tot$speaker== "adult"]) #6705
a_speaker_demuth*100/total_demuth -> d_perc_A#9.6%

a_speaker_french<-sum(sum_french_tot$nb_utt[sum_french_tot$speaker== "adult"]) #2137
a_speaker_french*100/total_french -> fr_perc_A #5.6%

m_speaker_demuth<-sum(sum_demuth_tot$nb_utt[sum_demuth_tot$speaker== "mother"]) #10283
m_speaker_demuth*100/total_demuth -> d_perc_M #14.8%

m_speaker_french<-sum(sum_french_tot$nb_utt[sum_french_tot$speaker== "mother"]) # 15706
m_speaker_french*100/total_french -> fr_perc_M #41%

peer_speaker_demuth<-sum(sum_demuth_tot$nb_utt[sum_demuth_tot$speaker== "other child"]) #18246
peer_speaker_demuth*100/total_demuth -> d_perc_PEER #26.2%

peer_speaker_french<-sum(sum_french_tot$nb_utt[sum_french_tot$speaker== "other child"]) #1833
peer_speaker_french*100/total_french -> fr_perc_PEER #4.8%

#SNAPSHOT
wholeset <- data.frame(
  corpus=c("Sesotho", "French","Sesotho", "French","Sesotho", "French","Sesotho", "French","Sesotho","French"), 
  production=c("Target Child","Target Child", "Adults","Adults", "Mother", "Mother", "Investigator", "Investigator","Other Children","Other Children"), 
  utt=c(tot_speaker_demuth, tot_speaker_french, a_speaker_demuth, a_speaker_french, m_speaker_demuth,m_speaker_french, inv_speaker_demuth, inv_speaker_french, peer_speaker_demuth, peer_speaker_french),
  value= c(d_perc_TC, fr_perc_TC,d_perc_A,fr_perc_A,d_perc_M,fr_perc_M,d_perc_INV,fr_perc_INV,d_perc_PEER,fr_perc_PEER))


Sesotho <- ggplot(subset(wholeset, corpus %in% c("Sesotho")), aes (x="", y = value, fill = production)) + geom_col(position = 'stack', width = 1,color="white") +geom_text(aes(label = paste(round(value,1), "%"), x = 1.3),position = position_stack(vjust = 0.5)) + theme_classic() +theme(plot.title = element_text(hjust=0.5),axis.line = element_blank(),axis.text = element_blank(),axis.ticks = element_blank()) + labs(fill = "Speaker", x = NULL,y = NULL,title = "Sesotho") + coord_polar("y")

French <- ggplot(subset(wholeset, corpus %in% c("French")), aes (x="", y = value, fill = production)) + geom_col(position = 'stack', width = 1,color="white") +geom_text(aes(label = paste(round(value,1), "%"), x = 1.3),position = position_stack(vjust = 0.5)) + theme_classic() +theme(plot.title = element_text(hjust=0.5),axis.line = element_blank(),axis.text = element_blank(),axis.ticks = element_blank()) + labs(fill = "Speaker", x = NULL,y = NULL,title = "French") + coord_polar("y")

ggarrange(Sesotho, French, ncol=2, nrow=1, common.legend = TRUE, legend="right")

```

## Distribution of speaker by corpus and keychild (INCLUDING INVESTIGATOR!!)
```{r cleaning and proportion, echo=F}

#demuth
dtot <- sum(tot_sum_demuth_register$nb_utt)#37843
h <- sum(tot_sum_demuth_register$nb_utt[tot_sum_demuth_register$keychild=="Hlobohang"]) #12634 33,3
ts <- sum(tot_sum_demuth_register$nb_utt[tot_sum_demuth_register$keychild=="Tsebo"]) #9883 26,1
l <- sum(tot_sum_demuth_register$nb_utt[tot_sum_demuth_register$keychild=="Litlhare"])# 15326 40,5

#french
ftot <-sum(tot_sum_french_register$nb_utt) #20917
anae <- sum(tot_sum_french_register$nb_utt[tot_sum_french_register$keychild=="Anae"]) #9677 46.2
anais <- sum(tot_sum_french_register$nb_utt[tot_sum_french_register$keychild=="Anais"])# 7212 34.5
th<- sum(tot_sum_french_register$nb_utt[tot_sum_french_register$keychild=="Theotime"]) #4028 19.2

#adding nbtot for proportion to the datasets
#demuth
tmp <- data.frame(
  keychild=c("Hlobohang","Tsebo","Litlhare"),
  tot_nb_utt=c(h,ts,l))
tot_sum_demuth_register_1 <- merge(tot_sum_demuth_register,tmp) 
#french
tmp2 <- data.frame(
  keychild=c("Anae","Anais","Theotime"),
  tot_nb_utt=c(anae,anais,th))
tot_sum_french_register_1 <- merge(tot_sum_french_register,tmp2) 

#creating a nb_utt controlled by total lenght by child
#demuth
aggregate(tot_sum_demuth_register_1$nb_utt,by=list(tot_sum_demuth_register_1$keychild, tot_sum_demuth_register_1$speaker,tot_sum_demuth_register_1$tot_nb_utt),sum)->d_speaker_prop
colnames(d_speaker_prop) <- c("keychild", "speaker", "tot_nb_utt", "nb_utt")

d_speaker_prop$prop_utt <- d_speaker_prop$nb_utt *100 / d_speaker_prop$tot_nb_utt

#french
aggregate(tot_sum_french_register_1$nb_utt,by=list(tot_sum_french_register_1$keychild, tot_sum_french_register_1$speaker,tot_sum_french_register_1$tot_nb_utt),sum)->fr_speaker_prop
colnames(fr_speaker_prop) <- c("keychild", "speaker", "tot_nb_utt", "nb_utt")

fr_speaker_prop$prop_utt <- fr_speaker_prop$nb_utt *100 / fr_speaker_prop$tot_nb_utt


```
## SNAPSHOT WITHOUT Target child 
````{r fig1, echo=F}
ggfillsesotho <-  ggplot(d_speaker_prop, aes( y=prop_utt, x="", fill=speaker))+geom_bar(stat="identity", position="fill")+  labs(x="Sesotho", y="Number of utterances") + guides(fill=FALSE) + scale_fill_discrete(labels= c("Adults", "Investigator", "Mother", "Other children")) + blank_theme                                                                      
ggfillfrench <- ggplot(fr_speaker_prop, aes( y=prop_utt, x="", fill=speaker))+geom_bar(stat="identity", position="fill")+  labs(x="Sesotho", y="Number of utterances")   + scale_fill_discrete(labels= c("Adults", "Investigator", "Mother", "Other children")) + blank_theme

ggarrange(ggfillsesotho, ggfillfrench, ncol=2, nrow=1, common.legend = TRUE, legend="right",  labels = c("Sesotho", "French"))

# dsummaryspeaker_tot<- data_summary(d_speaker_prop, varname="nb_utt", 
#                     groupnames=c("speaker", "keychild"))


ggrawsesotho <-  ggplot(d_speaker_prop, aes( y=as.numeric(prop_utt), x="", fill=speaker))+geom_bar(stat="identity", position="dodge")+  labs(x="Sesotho", y="Number of utterances") + guides(fill=FALSE) + scale_fill_discrete(labels= c("Adults", "Investigator", "Mother", "Other children")) + blank_theme 

ggrawfrench <- ggplot(fr_speaker_prop, aes( y=prop_utt, x="", fill=speaker))+geom_bar(stat="identity", position="dodge")+  labs(x="Sesotho", y="Number of utterances")   + scale_fill_discrete(labels= c("Adults", "Investigator", "Mother", "Other children")) + blank_theme

ggarrange(ggrawsesotho, ggrawfrench, ncol=2, nrow=1, common.legend = TRUE, legend="right",  labels = c("Sesotho", "French"))

kch_french <- ggplot(fr_speaker_prop, aes( y=nb_utt, x=keychild, fill=speaker))+geom_bar(stat="identity", position="dodge")+  labs(x="Keychild", y="Number of utterances")
kch_demuth <- ggplot(d_speaker_prop, aes( y=nb_utt, x=keychild, fill=speaker))+geom_bar(stat="identity", position="dodge")+  labs(x="Keychild", y="Number of utterances")+ guides(fill=FALSE)

ggarrange(kch_demuth, kch_french, ncol=2, nrow=1, common.legend = TRUE, legend="right", labels = c("Sesotho", "French"))

kchprop_french <- ggplot(fr_speaker_prop, aes( y=prop_utt, x=keychild, fill=speaker))+geom_bar(stat="identity", position="fill")+  labs(x="Keychild", y="Number of utterances") + scale_fill_discrete(labels= c("Adults", "Investigator", "Mother", "Other children")) + blank_theme
kchprop_demuth <- ggplot(d_speaker_prop, aes( y=prop_utt, x=keychild, fill=speaker))+geom_bar(stat="identity", position="fill")+  labs(x="Keychild", y="Number of utterances")+ guides(fill=FALSE) + scale_fill_discrete(labels= c("Adults", "Investigator", "Mother", "Other children")) + blank_theme

ggarrange(kchprop_demuth, kchprop_french, ncol=2, nrow=1, common.legend = TRUE, legend="right", labels = c("Sesotho", "French"))

```

## Distribution of speaker by corpus and keychild (EXCLUDE INVESTIGATOR!!)

```{r remove INV, echo=F}

#choose to remove Investigator // this is important for calculating percentages
#Remove investigator, unclassified speech
sum_demuth_register<-subset(tot_sum_demuth_register, !(tot_sum_demuth_register$speaker=="investigator"))
sum_demuth_register<-subset(sum_demuth_register, !(sum_demuth_register$input=="unclassified"))
sum_demuth_register<-subset(sum_demuth_register, !(sum_demuth_register$input=="Investigator"))

sum_french_register<-subset(tot_sum_french_register, !(tot_sum_french_register$speaker=="investigator"))
sum_french_register<-subset(sum_french_register, !(sum_french_register$input=="unclassified"))
sum_french_register<-subset(sum_french_register, !(sum_french_register$input=="Investigator"))

#demuth
dtot2 <- sum(sum_demuth_register$nb_utt)#31385
h2 <- sum(sum_demuth_register$nb_utt[sum_demuth_register$keychild=="Hlobohang"]) #11033 35.1
ts2 <- sum(sum_demuth_register$nb_utt[sum_demuth_register$keychild=="Tsebo"]) #7936 25.3
l2 <- sum(sum_demuth_register$nb_utt[sum_demuth_register$keychild=="Litlhare"])#  12416 39.6

#french
ftot2 <-sum(sum_french_register$nb_utt) #18075
anae2 <- sum(sum_french_register$nb_utt[sum_french_register$keychild=="Anae"]) #7120 39.4
anais2 <- sum(sum_french_register$nb_utt[sum_french_register$keychild=="Anais"])# 7002 38.7
th2<- sum(sum_french_register$nb_utt[sum_french_register$keychild=="Theotime"]) #3953 21.9

#adding nbtot for proportion to the datasets
#demuth
tmp3 <- data.frame(
  keychild=c("Hlobohang","Tsebo","Litlhare"),
  tot_nb_utt=c(h2,ts2,l2))
sum_demuth_register <- merge(sum_demuth_register,tmp3) 
#french
tmp4 <- data.frame(
  keychild=c("Anae","Anais","Theotime"),
  tot_nb_utt=c(anae2,anais2,th2))
sum_french_register <- merge(sum_french_register,tmp4) 

#creating a nb_utt controlled by total lenght by child and speaker only
#demuth
aggregate(sum_demuth_register$nb_utt,by=list(sum_demuth_register$keychild, sum_demuth_register$speaker,sum_demuth_register$tot_nb_utt),sum)->d_speaker_prop_clean
colnames(d_speaker_prop_clean) <- c("keychild", "speaker", "tot_nb_utt", "nb_utt")

d_speaker_prop_clean$prop_utt <- d_speaker_prop_clean$nb_utt *100 / d_speaker_prop_clean$tot_nb_utt

#french
aggregate(sum_french_register$nb_utt,by=list(sum_french_register$keychild, sum_french_register$speaker,sum_french_register$tot_nb_utt),sum)->fr_speaker_prop_clean
colnames(fr_speaker_prop_clean) <- c("keychild", "speaker", "tot_nb_utt", "nb_utt")

fr_speaker_prop_clean$prop_utt <- fr_speaker_prop_clean$nb_utt *100 / fr_speaker_prop_clean$tot_nb_utt

#creating a nb_utt controlled by total lenght by child and speaker and input type
#demuth
aggregate(sum_demuth_register$nb_utt,by=list(sum_demuth_register$keychild, sum_demuth_register$input, sum_demuth_register$speaker,sum_demuth_register$tot_nb_utt),sum)->d_speaker_prop_clean_1
colnames(d_speaker_prop_clean_1) <- c("keychild", "input","speaker", "tot_nb_utt", "nb_utt")

d_speaker_prop_clean_1$prop_utt <- d_speaker_prop_clean_1$nb_utt *100 / d_speaker_prop_clean_1$tot_nb_utt

sum_demuth_register <- merge(sum_demuth_register,d_speaker_prop_clean_1) 

#french
aggregate(sum_french_register$nb_utt,by=list(sum_french_register$keychild, sum_french_register$input, sum_french_register$speaker,sum_french_register$tot_nb_utt),sum)->fr_speaker_prop_clean_1
colnames(fr_speaker_prop_clean_1) <-c("keychild", "input","speaker", "tot_nb_utt", "nb_utt")

fr_speaker_prop_clean_1$prop_utt <- fr_speaker_prop_clean_1$nb_utt *100 / fr_speaker_prop_clean_1$tot_nb_utt

sum_french_register <- merge(sum_french_register,fr_speaker_prop_clean_1) 

```


Without investigator
```{r fig2 ,echo=F}
gg2fillsesotho <-  ggplot(d_speaker_prop_clean, aes( y=prop_utt, x="", fill=speaker))+geom_bar(stat="identity", position="fill")+  labs(x="Sesotho", y="Number of utterances") + guides(fill=FALSE) + scale_fill_discrete(labels= c("Adults", "Mother", "Other children")) + blank_theme

gg2fillfrench <- ggplot(fr_speaker_prop_clean, aes( y=prop_utt, x="", fill=speaker))+geom_bar(stat="identity", position="fill")+  labs(x="Sesotho", y="Number of utterances")   + scale_fill_discrete(labels= c("Adults", "Mother", "Other children")) + blank_theme

ggarrange(gg2fillsesotho, gg2fillfrench, ncol=2, nrow=1, common.legend = TRUE, legend="right",  labels = c("Sesotho", "French"))

gg2rawsesotho <-  ggplot(d_speaker_prop_clean, aes( y=prop_utt, x="", fill=speaker))+geom_bar(stat="identity", position="dodge")+  labs(x="Sesotho", y="Number of utterances") + guides(fill=FALSE) + scale_fill_discrete(labels= c("Adults","Mother", "Other children")) + blank_theme

gg2rawfrench <- ggplot(fr_speaker_prop_clean, aes( y=prop_utt, x="", fill=speaker))+geom_bar(stat="identity", position="dodge")+  labs(x="Sesotho", y="Number of utterances")   + scale_fill_discrete(labels= c("Adults", "Mother", "Other children")) + blank_theme

ggarrange(gg2rawsesotho, gg2rawfrench, ncol=2, nrow=1, common.legend = TRUE, legend="right",  labels = c("Sesotho", "French"))

kch_french <- ggplot(fr_speaker_prop_clean, aes( y=nb_utt, x=keychild, fill=speaker))+geom_bar(stat="identity", position="dodge")+  labs(x="Keychild", y="Number of utterances")  + scale_fill_discrete(labels= c("Adults", "Mother", "Other children")) + blank_theme
kch_demuth <- ggplot(d_speaker_prop_clean, aes( y=nb_utt, x=keychild, fill=speaker))+geom_bar(stat="identity", position="dodge")+  labs(x="Keychild", y="Number of utterances")+ guides(fill=FALSE)  + scale_fill_discrete(labels= c("Adults", "Mother", "Other children")) + blank_theme

ggarrange(kch_demuth, kch_french, ncol=2, nrow=1, common.legend = TRUE, legend="right", labels = c("Sesotho", "French"))

kchprop_french <- ggplot(fr_speaker_prop_clean, aes( y=prop_utt, x=keychild, fill=speaker))+geom_bar(stat="identity", position="fill")+  labs(x="Keychild", y="Number of utterances") + scale_fill_discrete(labels= c("Adults", "Mother", "Other children")) + blank_theme
kchprop_demuth <- ggplot(d_speaker_prop_clean, aes( y=prop_utt, x=keychild, fill=speaker))+geom_bar(stat="identity", position="fill")+  labs(x="Keychild", y="Number of utterances")+ guides(fill=FALSE)+ scale_fill_discrete(labels= c("Adults", "Mother", "Other children")) + blank_theme

ggarrange(kchprop_demuth, kchprop_french, ncol=2, nrow=1, common.legend = TRUE, legend="right", labels = c("Sesotho", "French"))
```

```{r cds and ocds, echo=F, message=FALSE}
#combine input ADS+OCDS
#demuth
sum_demuth_register$type_speech[ sum_demuth_register$input %in% c("ADS","OCDS") ] <- "Overheard" 
sum_demuth_register$type_speech[ sum_demuth_register$input %in% c("CDS") ] <- "Directed"

#french
sum_french_register$type_speech[ sum_french_register$input %in% c("ADS","OCDS") ] <- "Overheard" 
sum_french_register$type_speech[ sum_french_register$input %in% c("CDS") ] <- "Directed"

summary_d<- summarySE(sum_demuth_register, measurevar="prop_utt", 
                    groupvars=c("speaker", "type_speech"))
summary_f<- summarySE(sum_french_register, measurevar="prop_utt", 
                    groupvars=c("speaker", "type_speech"))

summary_d_lara <-summary_d<- summarySE(sum_demuth_register, measurevar="prop_utt", 
                    groupvars=c("speaker", "input"))
summary_f_lara<- summarySE(sum_french_register, measurevar="prop_utt", 
                    groupvars=c("speaker", "input"))


```

## Input composition
```{r percentages, echo=F, message=F, results='hide'}

sum(sum_french_register$prop_utt[sum_french_register$type_speech=="Directed"]) /3 #91.6%
sum(sum_demuth_register$prop_utt[sum_demuth_register$type_speech=="Directed"]) /3 #79.6%

sum(sum_french_register$prop_utt[sum_french_register$type_speech=="Directed" & sum_french_register$speaker =="mother"]) /3 #79.1%
sum(sum_demuth_register$prop_utt[sum_demuth_register$type_speech=="Directed" & sum_demuth_register$speaker =="mother"]) /3 #25.3%

sum(sum_french_register$prop_utt[sum_french_register$type_speech=="Directed" & sum_french_register$speaker =="other child"]) /3 #3.3%
sum(sum_demuth_register$prop_utt[sum_demuth_register$type_speech=="Directed" & sum_demuth_register$speaker =="other child"]) /3 #38.5%

sum(sum_french_register$prop_utt[sum_french_register$type_speech=="Directed" & sum_french_register$speaker =="adult"]) /3 #9.2%
sum(sum_demuth_register$prop_utt[sum_demuth_register$type_speech=="Directed" & sum_demuth_register$speaker =="adult"]) /3 #15.8%

```
CDS is `r round(sum(sum_french_register$prop_utt[sum_french_register$type_speech=="Directed"]) /3,1) `% for French and `r round(sum(sum_demuth_register$prop_utt[sum_demuth_register$type_speech=="Directed"]) /3,1) `% for Sesotho.


In terms of CDS input composition:

- For French
It is on average `r round(sum(sum_french_register$prop_utt[sum_french_register$type_speech=="Directed" & sum_french_register$speaker =="mother"]) /3,1)`% from the mother; `r round(sum(sum_french_register$prop_utt[sum_french_register$type_speech=="Directed" & sum_french_register$speaker =="adult"]) /3,1)`% from adults and `r round(sum(sum_french_register$prop_utt[sum_french_register$type_speech=="Directed" & sum_french_register$speaker =="other child"]) /3,1)`% from other children.


- For Sesotho: It is on average `r round(sum(sum_demuth_register$prop_utt[sum_demuth_register$type_speech=="Directed" & sum_demuth_register$speaker =="mother"]) /3,1)`% from the mother; `r round(sum(sum_demuth_register$prop_utt[sum_demuth_register$type_speech=="Directed" & sum_demuth_register$speaker =="adult"]) /3,1)`% from adults and `r round(sum(sum_demuth_register$prop_utt[sum_demuth_register$type_speech=="Directed" & sum_demuth_register$speaker =="other child"]) /3,1)`% from other children.


```{r fig3, echo=F,message=FALSE, results='hide'}
goodgraph_sesotho<- ggplot(summary_d, aes(x=type_speech, y=prop_utt, fill=speaker)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black", size=0.3) +      # Thinner lines
    geom_errorbar(aes(ymin=prop_utt-se, ymax=prop_utt+se),
                  size=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
    xlab("Sesotho") +
    ylab("Proportion of mean utterance count") +
    scale_fill_hue(name="Speaker", # Legend label, use darker colors
                   breaks=c("adult", "mother", "other child"),
                   labels=c("Adults", "Mother", "Other children")) + scale_y_continuous(breaks=0:80*5) + ylim(0, 90)
    theme_bw()

goodgraph_french<- ggplot(summary_f, aes(x=type_speech, y=prop_utt, fill=speaker)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black", size=0.3) +      # Thinner lines
    geom_errorbar(aes(ymin=prop_utt-se, ymax=prop_utt+se),
                  size=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
    xlab("French") +
    ylab("") +
    scale_fill_hue(name="Speaker", # Legend label, use darker colors
                   breaks=c("adult", "mother", "other child"),
                   labels=c("Adults", "Mother", "Other children")) +  scale_y_continuous(breaks=0:80*5) + ylim(0, 90)
    theme_bw()

forBU <- ggarrange(goodgraph_sesotho, goodgraph_french, ncol=2, nrow=1, common.legend = TRUE, legend="right")
forBU
ggsave("forBU.png")

## If you want other color palette

goodgraph_sesotho<- ggplot(summary_d, aes(x=type_speech, y=prop_utt, fill=speaker)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black") +      # Thinner lines
    geom_errorbar(aes(ymin=prop_utt-se, ymax=prop_utt+se),
                  size=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
    xlab("Sesotho") +
    ylab("Proportion of mean utterance count") +
    scale_fill_manual(name="Speaker", # Legend label, use darker colors
                   breaks=c("adult", "mother", "other child"),
                   labels=c("Adults", "Mother", "Other children"), values=c("#caf2f1", "#6abac3","#beb9bd")) + scale_y_continuous(breaks=0:80*5) + ylim(0, 90)
    theme_bw()

goodgraph_french<- ggplot(summary_f, aes(x=type_speech, y=prop_utt, fill=speaker)) + 
    geom_bar(position=position_dodge(), stat="identity",
             colour="black") +      # Thinner lines
    geom_errorbar(aes(ymin=prop_utt-se, ymax=prop_utt+se),
                  size=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.9)) +
    xlab("French") +
    ylab("") +
    scale_fill_manual(name="Speaker", # Legend label, use darker colors
                   breaks=c("adult", "mother", "other child"),
                   labels=c("Adults", "Mother", "Other children"), values=c("#caf2f1", "#6abac3","#beb9bd")) +  scale_y_continuous(breaks=0:80*5) + ylim(0, 90)
    theme_bw()

ggarrange(goodgraph_sesotho, goodgraph_french, ncol=2, nrow=1, common.legend = TRUE, legend="right")



```


```{r demuth for figure purposes, echo=F}
######checked up to here -g
library(reshape2)
# #add label of type of input
# sum_demuth_ADS$input <- "ADS"
# sum_demuth_CDS$input <- "CDS"
# 
# #remove for CDS the target child utterances in which he is talking to himself
# sum_demuth_CDS<-subset(sum_demuth_CDS, !(sum_demuth_CDS$speaker=="Target_Child"))
# sum_demuth_ADS<-subset(sum_demuth_ADS, !(sum_demuth_ADS$speaker=="Target_Child"))
# 
# 
# #make sure that all children have the 4 categories ,if not mark as 0
# #CDS
# table(sum_demuth_CDS$keychild,sum_demuth_CDS$speaker) #check
# #create the lines (this should be done via a loop sorry!)
# peer_Hlo <- data.frame("sibling","Hlobohang", "0", "CDS")
# #names the lines
# names(peer_Hlo)<-c("speaker","keychild", "nb_utt","input")
# neww <- rbind(sum_demuth_CDS, peer_Hlo)
# #added to the main data.frame
# sum_demuth_CDS <- rbind(sum_demuth_CDS, peer_Hlo)
# 
# 
# #ADS
# table(sum_demuth_ADS$keychild,sum_demuth_ADS$speaker) #check
# #create the lines (this should be done via a loop sorry!)
# peer_Hlo <- data.frame("sibling","Hlobohang", "0", "ADS")
# #names the lines
# names(peer_Hlo)<-c("speaker","keychild", "nb_utt","input")
# neww <- rbind(sum_demuth_ADS, peer_Hlo)
# #added to the main data.frame
# sum_demuth_ADS <- rbind(sum_demuth_ADS, peer_Hlo)
# 
# demuth_fig <-rbind(sum_demuth_ADS,sum_demuth_CDS)
# trywells <- melt(sum_wells_ADS[,c("child","tot.adults", "tot.children")],id.vars = 1)
# colnames(dfm) <- c("child", "input", "value")
# dft <- melt(bigsum_low[,c("child","cds.adults", "cds.children")],id.vars = 1)
# colnames(dft) <- c("child", "input.cds", "value.cds")
# cbind(dfm,dft) -> try
# try[4] <- NULL
# merge(x=bigsum_low,y = try[ , c("child",  "input", "value", "input.cds", "value.cds"),], all.x=T,all.y=F)->try2
```


```{r figure making, echo=F, eval=F, include=TRUE, message=FALSE}
# 
# #not normalized / all corpus
# 
# 
# 
# 
# 
# 
# ggplot(sum_demuth_register, aes( y=nb_utt, x=keychild, fill=input))+ geom_bar(stat="identity", position="dodge")
# 
# Hlobohang <- ggplot(subset(sum_demuth_register_fig, keychild %in% c("Hlobohang") ), aes( y=nb_utt, x=input, fill=speaker))+
#   geom_bar(stat="identity", position="dodge") +  labs(x="Hlobohang", y="")  + guides(fill=FALSE)
# 
# Litlhare <- ggplot(subset(sum_demuth_register_fig, keychild %in% c("Litlhare") ), aes( y=nb_utt, x=input, fill=speaker))+
#   geom_bar(stat="identity", position="dodge") +  labs(x="Litlhare", y="Number of utterances")  + guides(fill=guide_legend(title=NULL))
# 
# Tsebo <- ggplot(subset(sum_demuth_register_fig, keychild %in% c("Tsebo") ), aes( y=nb_utt, x=input, fill=speaker))+
#   geom_bar(stat="identity", position="dodge") +  labs(x="Tsebo", y="")  + guides(fill=FALSE)
# 
# 
# Hlobohang_stack<- ggplot(subset(sum_demuth_register, keychild %in% c("Hlobohang") ), aes( y=nb_utt, x=input, fill=speaker))+
#   geom_bar(stat="identity", position="fill") +  labs(x="Hlobohang", y="")  + guides(fill=guide_legend(title=NULL))
# 
# Litlhare_stack<- ggplot(subset(sum_demuth_register, keychild %in% c("Litlhare") ), aes( y=nb_utt, x=input, fill=speaker))+
#   geom_bar(stat="identity", position="fill") +  labs(x="Litlhare", y="Number of utterances")  + guides(fill=guide_legend(title=NULL))
# 
# Tsebo_stack <- ggplot(subset(sum_demuth_register, keychild %in% c("Tsebo") ), aes( y=nb_utt, x=input, fill=speaker))+
#   geom_bar(stat="identity", position="fill") +  labs(x="Tsebo", y="")  + guides(fill=FALSE)
# #scale_fill_discrete(labels= c("Adults", "Mother", "Peers", "Siblings"))
# 
# #this plot for abstract
# Sesothofig<-ggplot(sum_demuth_register_fig, aes( y=nb_utt, x=input, fill=speaker))+
#   geom_bar(stat="identity", position="dodge") +  labs(x="", y="")  + guides(fill=FALSE) + ggtitle("Sesotho")
# Frenchfig<-ggplot(sum_french_register_fig, aes( y=nb_utt, x=input, fill=speaker))+
#   geom_bar(stat="identity", position="dodge") +  labs(x="", y="")  + guides(fill=guide_legend(title=NULL)) + ggtitle("French")
# 
# png(filename="/Users/admin/Desktop/name.png", res=100, width=500)
# ggarrange(Sesothofig, Frenchfig, ncol=1, nrow=2, common.legend = TRUE, legend="bottom")
# dev.off()
# 
# 
# #this
# ggarrange(Hlobohang, Litlhare,Tsebo, ncol=1, nrow=3, common.legend = TRUE, legend="bottom")
# ggarrange(Hlobohang_stack, Litlhare_stack,Tsebo_stack, ncol=1, nrow=3, common.legend = TRUE, legend="bottom")
# 
# 
# #panel
# multiplot (wells_full, demuth_full, cols=1)
# multiplot (wells_annot, demuth_annot, cols=1)
#  

#fill <- multiplot (wells_CDS, demuth_CDS, cols=1) 
#dodge <- multiplot (Hlobohang, Litlhare,Tsebo, cols=1) 
```
